% ============================================================
%  EthioSME Valuation System — Production Hosting &
%  Deployment Strategy
%  Prepared: 2026-06-23
% ============================================================
\documentclass[12pt,a4paper]{report}

% ── Packages ────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{mdframed}
\usepackage{tcolorbox}
\usepackage{float}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}

% ── Colours ─────────────────────────────────────────────────
\definecolor{primaryblue}{HTML}{1E3A5F}
\definecolor{accentblue}{HTML}{2E6DA4}
\definecolor{lightgray}{HTML}{F5F5F5}
\definecolor{darkgray}{HTML}{333333}
\definecolor{successgreen}{HTML}{2E7D32}
\definecolor{warningorange}{HTML}{E65100}
\definecolor{dangerred}{HTML}{C62828}
\definecolor{codebg}{HTML}{F8F8F8}
\definecolor{codefg}{HTML}{24292E}

% ── Hyperref Setup ──────────────────────────────────────────
\hypersetup{
    colorlinks=true,
    linkcolor=accentblue,
    urlcolor=accentblue,
    citecolor=accentblue,
    pdfauthor={EthioSME Thesis},
    pdftitle={Production Hosting and Deployment Strategy},
    pdfsubject={DevOps Architecture},
}

% ── Listings (code blocks) ──────────────────────────────────
\lstdefinestyle{shell}{
    backgroundcolor=\color{codebg},
    basicstyle=\ttfamily\footnotesize\color{codefg},
    breaklines=true,
    breakatwhitespace=true,
    frame=single,
    framesep=4pt,
    rulecolor=\color{lightgray},
    xleftmargin=8pt,
    xrightmargin=8pt,
    captionpos=b,
    numbers=none,
    showstringspaces=false,
}

\lstdefinestyle{yaml}{
    style=shell,
    morekeywords={true,false,null},
    keywordstyle=\color{accentblue}\bfseries,
}

\lstset{style=shell}

% ── Section Formatting ──────────────────────────────────────
\titleformat{\chapter}[display]
    {\normalfont\Large\bfseries\color{primaryblue}}
    {\chaptertitlename\ \thechapter}{10pt}{\Huge}
\titleformat{\section}
    {\normalfont\Large\bfseries\color{primaryblue}}
    {\thesection}{1em}{}
\titleformat{\subsection}
    {\normalfont\large\bfseries\color{accentblue}}
    {\thesubsection}{1em}{}
\titleformat{\subsubsection}
    {\normalfont\normalsize\bfseries\color{darkgray}}
    {\thesubsubsection}{1em}{}

% ── Header / Footer ─────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\color{accentblue}EthioSME Valuation System}
\fancyhead[R]{\small\color{accentblue}Deployment Strategy}
\fancyfoot[C]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\headrule}{\hbox to\headwidth{\color{accentblue}\leaders\hrule height \headrulewidth\hfill}}

% ── Info Boxes ──────────────────────────────────────────────
\tcbuselibrary{skins,breakable}

\newtcolorbox{infobox}[1][]{
    colback=accentblue!8,
    colframe=accentblue,
    fonttitle=\bfseries,
    title=#1,
    breakable,
    before upper={\parindent0pt},
}

\newtcolorbox{warningbox}[1][]{
    colback=warningorange!10,
    colframe=warningorange,
    fonttitle=\bfseries,
    title=#1,
    breakable,
}

\newtcolorbox{dangerbox}[1][]{
    colback=dangerred!8,
    colframe=dangerred,
    fonttitle=\bfseries,
    title=#1,
    breakable,
}

\newtcolorbox{successbox}[1][]{
    colback=successgreen!8,
    colframe=successgreen,
    fonttitle=\bfseries,
    title=#1,
    breakable,
}

% ── Column Types ────────────────────────────────────────────
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

% ── Document ────────────────────────────────────────────────
\begin{document}

% ── Title Page ──────────────────────────────────────────────
\begin{titlepage}
    \centering
    \vspace*{2cm}

    {\color{primaryblue}\rule{\linewidth}{2pt}}
    \vspace{0.5cm}

    {\Huge\bfseries\color{primaryblue}
        EthioSME Valuation System\\[0.4em]
        Production Hosting \&\\[0.2em]
        Deployment Strategy
    }

    \vspace{0.5cm}
    {\color{primaryblue}\rule{\linewidth}{2pt}}

    \vspace{1.5cm}

    {\large\color{darkgray}Senior DevOps Architecture Analysis}

    \vspace{0.8cm}

    \begin{tabular}{rl}
        \textbf{Prepared:}  & 23 June 2026 \\[4pt]
        \textbf{Scope:}     & Laravel 12 + React/Inertia.js Backend \\[2pt]
                            & Python FastAPI AI Service \\[2pt]
                            & Supabase PostgreSQL \\[2pt]
                            & HuggingFace Model Storage \\[4pt]
        \textbf{Version:}   & 1.0.0 \\
    \end{tabular}

    \vfill

    {\small\color{darkgray}
        This document constitutes the complete production deployment
        strategy for the EthioSME Valuation System thesis project.
        It covers architecture, infrastructure, security, CI/CD,
        monitoring, backup, cost estimation, and a deployment roadmap.
    }

    \vspace{1cm}
    {\color{primaryblue}\rule{\linewidth}{1pt}}
\end{titlepage}

% ── Table of Contents ───────────────────────────────────────
\tableofcontents
\newpage

% ════════════════════════════════════════════════════════════
\chapter{Baseline Reality Check}
\label{ch:baseline}
% ════════════════════════════════════════════════════════════

Before any recommendations, it is essential to understand the exact
current state of the system. The following analysis is derived from
direct inspection of both codebases.

\section{Current System Inventory}

\begin{longtable}{L{4cm} L{5.5cm} C{2.5cm}}
    \toprule
    \textbf{Concern} & \textbf{Current State} & \textbf{Production-Ready?} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Concern} & \textbf{Current State} & \textbf{Production-Ready?} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Database &
    Supabase PostgreSQL, SSL configured &
    \textcolor{successgreen}{\textbf{Yes}} \\[4pt]

    AI Service &
    HuggingFace Spaces (\texttt{leykun-code-...hf.space}) &
    \textcolor{warningorange}{\textbf{Partial}} \\[4pt]

    Laravel Dockerfile &
    Sail dev image, Vite dev server running in container &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    Queue Driver &
    \texttt{sync} — blocking, no async workers &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    File Storage &
    Local disk (ephemeral in containers) &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    Session / Cache &
    File-based drivers &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    SSL / HTTPS &
    Not configured on application side &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    CI/CD Pipeline &
    No pipeline exists &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    Background Jobs &
    No \texttt{app/Jobs/} directory &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    Reverse Proxy &
    None configured &
    \textcolor{dangerred}{\textbf{No}} \\[4pt]

    Health Checks &
    AI service has \texttt{GET /health}; Laravel does not &
    \textcolor{warningorange}{\textbf{Partial}} \\

\end{longtable}

\begin{infobox}[Key Finding]
    Two of four core infrastructure concerns — the database (Supabase)
    and the AI service (HuggingFace Spaces) — are already deployed to
    the cloud. The critical gap is a production-ready Laravel
    application deployment with correct process management, storage,
    session handling, and build pipeline.
\end{infobox}

\section{System Overview}

\subsection{Laravel Backend}

\begin{itemize}[leftmargin=1.5em]
    \item Full-stack PHP 8.3 application with Domain-Driven Design
    \item React 18 frontend embedded via Inertia.js (single-origin monolith)
    \item 21 models, 10 domain modules, 41 controllers
    \item JWT authentication (API) + session-based (web/Inertia)
    \item PostgreSQL database via Supabase
    \item Audit logging via \texttt{owen-it/laravel-auditing}
    \item No background job infrastructure (sync queue driver)
\end{itemize}

\subsection{Python AI Service}

\begin{itemize}[leftmargin=1.5em]
    \item FastAPI microservice for cashflow forecasting and credit scoring
    \item Three ML models: DeepAR (probabilistic forecasting),
          XGBoost + SHAP (credit scoring), optional LSTM encoder
    \item Shared PostgreSQL connection with Laravel (feature SQL queries)
    \item Model artifact persistence via HuggingFace Hub
    \item Structured JSON logging via \texttt{structlog}
    \item Already deployed at HuggingFace Spaces
\end{itemize}

\subsection{Loan Application State Machine}

\begin{center}
\small
\texttt{draft → submitted → pending\_psychometric → pending\_data\_sync}\\
\texttt{→ queued\_for\_ai → processing → evaluated → \{approved | rejected | withdrawn\}}
\end{center}

\subsection{Contract Versions}

Both services are pinned to \texttt{AI\_SERVICE\_CONTRACT\_VERSION=v1}.
The integration endpoint is \texttt{POST /predict} on the Python service,
called synchronously by Laravel's \texttt{ValuationService}.

% ════════════════════════════════════════════════════════════
\chapter{Production Architecture Proposal}
\label{ch:architecture}
% ════════════════════════════════════════════════════════════

\section{Recommended Architecture}

The recommended approach is a \textbf{cloud-native managed services}
architecture, optimised for thesis demonstration requirements: low cost,
high reliability for demo sessions, minimal operational overhead, and
no over-engineering beyond what the project requires.

\subsection{Architecture Diagram}

\begin{lstlisting}[caption={High-Level Production Architecture}]
+------------------------------------------------------------------+
|                         CLOUDFLARE                               |
|             DNS + CDN + WAF + SSL Termination                   |
|         (Free tier -- covers all traffic, HTTPS globally)       |
+----------------------------+---------------------------------+---+
                             | HTTPS (443)
                             v
+------------------------------------------------------------------+
|                    RAILWAY (Primary Host)                        |
|              Laravel 12 + React/Inertia.js                      |
|          PHP 8.3 + PHP-FPM + Nginx (production image)           |
|                                                                  |
|   +---------------+   +--------------+   +------------------+   |
|   |  Web Process  |   | Queue Worker |   |   Scheduler      |   |
|   |  (php-fpm +   |   | (queue:work) |   | (artisan:        |   |
|   |   nginx)      |   |              |   |  schedule:run)   |   |
|   +---------------+   +--------------+   +------------------+   |
+-------------+--------------------------------------------+------+
              |                                            |
      +-------+------+  +------------------+  +-----------+--------+
      |   Supabase   |  |  HuggingFace     |  |  Cloudflare R2     |
      |  PostgreSQL  |  |  Spaces (AI)     |  |  (Object Storage)  |
      |              |  |  FastAPI +       |  |                    |
      |  Primary DB  |  |  XGBoost +       |  |  File uploads,     |
      |  (shared by  |  |  DeepAR + LSTM   |  |  Excel exports     |
      |  both svcs)  |  |                  |  |                    |
      +--------------+  +------------------+  +--------------------+
                               |
                        Shared PostgreSQL
                        (same Supabase DB)
\end{lstlisting}

\subsection{Service Responsibilities}

\begin{longtable}{L{3.5cm} L{3.5cm} L{6cm}}
    \toprule
    \textbf{Service} & \textbf{Host} & \textbf{Role} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Service} & \textbf{Host} & \textbf{Role} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Laravel app (web) & Railway / Fly.io &
        HTTP request handling, Inertia.js rendering, REST API routes \\[4pt]

    Laravel queue worker & Railway (separate process) &
        Background job processing (valuation callbacks, email) \\[4pt]

    Laravel scheduler & Railway (cron process) &
        Recurring tasks, keep-alive pings, macroeconomic ingestion \\[4pt]

    FastAPI AI service & HuggingFace Spaces &
        Valuation endpoint, model training, drift metrics \\[4pt]

    PostgreSQL & Supabase &
        Primary database shared by both services \\[4pt]

    Object storage & Cloudflare R2 &
        Excel reports, any uploaded files \\[4pt]

    CDN + SSL + DNS & Cloudflare &
        HTTPS everywhere, static asset caching, DDoS protection \\[4pt]

    Error tracking & Sentry (free) &
        Exception capture, performance tracing, alerting \\[4pt]

    Uptime monitoring & UptimeRobot (free) &
        Availability alerts, endpoint health checks \\

\end{longtable}

\subsection{Data Flow}

\begin{lstlisting}[caption={Request Flow: User to AI Valuation}]
User (Browser)
    --> Cloudflare CDN (SSL termination, cache static assets)
        --> Railway: Laravel (PHP-FPM + Nginx, port 80)
            --> Supabase: PostgreSQL (read/write application data)
            --> HuggingFace Spaces: FastAPI
                    POST /predict (X-Internal-Key header)
                --> Supabase: PostgreSQL
                        (read heartbeat, psychometric, exogenous data)
                --> ML Models (/data/models/ on HF Spaces)
                        (XGBoost + DeepAR + SHAP computation)
                <-- PredictResponse (NPV, credit limit, SHAP drivers)
            --> Supabase: PostgreSQL (persist Valuation + ShapExplanation)
        <-- Inertia response / JSON
    <-- HTTPS response
\end{lstlisting}

% ════════════════════════════════════════════════════════════
\chapter{Infrastructure Requirements}
\label{ch:infrastructure}
% ════════════════════════════════════════════════════════════

\section{What Must Change Before Deployment}

\subsection{Critical Blockers}

\subsubsection{Replace the Sail Development Dockerfile}

The current \texttt{Dockerfile} uses \texttt{sail-8.5/app:latest} as
the base image. This is a development image that includes Xdebug,
development tooling, and is not hardened for production. Additionally,
the \texttt{docker-entrypoint.sh} starts both the Vite development
server and the PHP built-in server (\texttt{php artisan serve}).
Neither is suitable for production:

\begin{itemize}[leftmargin=1.5em]
    \item The PHP built-in server is single-threaded and designed only
          for local development.
    \item The Vite development server performs hot module replacement
          (HMR) — a development-only feature.
\end{itemize}

A production image must use \texttt{php:8.3-fpm-alpine} with
Nginx as the web server front-end.

\subsubsection{Frontend Assets Must Be Pre-Built}

In production, \texttt{npm run build} must compile all TypeScript/React
assets to \texttt{public/build/} during the Docker image build stage.
Nginx then serves these as static files. The Vite development server
must not run in any production process.

\subsubsection{Queue Driver Must Change from \texttt{sync}}

The current \texttt{QUEUE\_CONNECTION=sync} means all queued work
blocks the web request and runs synchronously. The valuation pipeline
— AI service HTTP calls, SHAP storage, audit logging — is already
computationally heavy. In production, the queue must use the
\texttt{database} driver with a persistent worker process running
\texttt{php artisan queue:work}.

\subsubsection{Session and Cache Must Not Use the File Driver}

File-based sessions (\texttt{SESSION\_DRIVER=file}) and file-based
cache (\texttt{CACHE\_STORE=file}) do not work correctly across
multiple container instances and are slow under concurrency. Production
must use the \texttt{database} driver for both.

\subsection{Important Pre-Launch Items}

\subsubsection{Laravel Health Check Endpoint}

The Python AI service exposes \texttt{GET /health}. Laravel has no
equivalent. A \texttt{/up} route is required for:

\begin{itemize}[leftmargin=1.5em]
    \item Railway health checks (to determine when a deploy is healthy)
    \item Cloudflare origin health checks
    \item UptimeRobot monitoring
\end{itemize}

Laravel 11+ includes a default \texttt{/up} route via
\texttt{Illuminate\textbackslash Foundation\textbackslash Http\textbackslash Middleware}.
Verify this is active in the production routing.

\subsubsection{Storage Driver Must Change to S3-Compatible}

Any files written to the \texttt{local} disk are lost on container
restart (containers use ephemeral storage). The S3 driver is already
configured in \texttt{config/filesystems.php} but unused. Cloudflare R2
provides an S3-compatible API with no egress fees and a free tier of
10\,GB/month.

\subsubsection{Logging Must Go to Standard Output}

In containerised environments, log files in \texttt{storage/logs/} are
ephemeral and invisible to the platform log viewer. Laravel must be
configured to write logs to \texttt{stderr} using the \texttt{stderr}
channel in \texttt{config/logging.php}.

\subsubsection{APP\_KEY and Secrets Management}

\texttt{APP\_KEY}, \texttt{JWT\_SECRET}, \texttt{AI\_SERVICE\_TOKEN},
and all other secrets must be injected as platform environment variable
secrets (Railway secret store). They must never appear in any committed
file.

\subsubsection{HuggingFace Spaces Cold Start}

The HuggingFace Spaces free tier suspends spaces after 48 hours of
inactivity. The Laravel \texttt{AI\_SERVICE\_FALLBACK\_ENABLED=true}
flag provides degraded mode, but a sleeping Space means the core
valuation feature is unavailable. A keep-alive cron job that pings
the \texttt{/health} endpoint every 30 minutes is essential.

% ════════════════════════════════════════════════════════════
\chapter{Technology Stack Recommendations}
\label{ch:techstack}
% ════════════════════════════════════════════════════════════

\section{Laravel Hosting: Railway}

\begin{warningbox}[Primary Recommendation: Railway.app]
    Railway is the recommended host for the Laravel application.
    It supports multi-process Docker deployments, never sleeps, provides
    built-in environment variable secrets, and costs \$0--5/month at
    thesis scale.
\end{warningbox}

\subsection{Comparison of Laravel Hosting Options}

\begin{longtable}{L{2.8cm} C{2cm} L{7.5cm}}
    \toprule
    \textbf{Platform} & \textbf{Verdict} & \textbf{Rationale} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Platform} & \textbf{Verdict} & \textbf{Rationale} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    \textbf{Railway} &
    \textcolor{successgreen}{\textbf{Recommended}} &
    Docker-native; multiple processes per project (web, worker,
    scheduler); \$5/mo credits; never sleeps; environment variable
    management; GitHub auto-deploy \\[6pt]

    Fly.io &
    \textcolor{accentblue}{Alternative} &
    More control; generous free tier; global edge; but requires
    \texttt{fly.toml} and additional CLI configuration \\[6pt]

    Render.com &
    \textcolor{warningorange}{Acceptable} &
    Docker deploy works, but free tier spins down after 15 minutes
    --- fatal for a thesis demo \\[6pt]

    Vercel &
    \textcolor{warningorange}{Possible but wrong fit} &
    Serverless PHP has critical limitations: no persistent queue
    workers, no storage, cold starts, 10-second function timeout;
    despite \texttt{laravel-vercel-installer} being present, this
    architecture is too stateful for serverless \\[6pt]

    DigitalOcean App Platform &
    \textcolor{accentblue}{Solid} &
    \$12/mo, managed containers, good CLI, slightly more expensive \\[6pt]

    AWS ECS / Fargate &
    \textcolor{dangerred}{Overkill} &
    Correct tool for enterprise scale; excessive operational overhead
    for a single-developer thesis project \\[6pt]

    Self-managed VPS &
    \textcolor{warningorange}{Possible} &
    \$6/mo, full control, but requires manual Nginx/SSL/deployment
    setup and ongoing maintenance \\

\end{longtable}

\subsection{Why Railway is Correct for This Project}

\begin{enumerate}[leftmargin=1.5em]
    \item One \texttt{railway up} deployment from a Dockerfile
    \item Supports multiple services (web, worker, cron) in one project
    \item Built-in environment variable secrets management
    \item Auto-deploys on \texttt{git push} via GitHub integration
    \item \$5/month free credits sufficient for thesis traffic
    \item Persistent storage volumes available if needed
    \item Services never sleep between requests
    \item Web dashboard for log viewing and rollbacks
\end{enumerate}

\section{Database: Supabase PostgreSQL (Keep Existing)}

The database is already configured, connected, and working. Supabase
provides the following without any additional setup:

\begin{itemize}[leftmargin=1.5em]
    \item Connection pooling via PgBouncer (critical for PHP's
          connection-per-request model)
    \item SSL enforced (\texttt{DB\_SSLMODE=require} already set)
    \item Daily automated backups (7-day retention on free tier)
    \item Row Level Security for multi-tenancy
    \item Free tier: 500\,MB storage, 5\,GB egress
\end{itemize}

\begin{warningbox}[Supabase Free Tier Risk]
    The Supabase free tier pauses the database after 7 days of
    inactivity. For demo reliability, either upgrade to Supabase Pro
    (\$25/mo) or implement a database keep-alive cron job.
\end{warningbox}

\section{AI Service: HuggingFace Spaces (Persistent Tier)}

The FastAPI service is already deployed and the Dockerfile is already
HuggingFace-compatible (port 7860, non-root user, Python 3.11 slim).
The model artifact storage via HuggingFace Hub is already integrated in
\texttt{storage\_service.py}.

\begin{longtable}{L{3cm} L{2.5cm} L{2.5cm} L{4cm}}
    \toprule
    \textbf{HF Tier} & \textbf{CPU} & \textbf{RAM} & \textbf{Behaviour} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{HF Tier} & \textbf{CPU} & \textbf{RAM} & \textbf{Behaviour} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Free (CPU) & 2 vCPU & 16\,GB & Sleeps after 48h inactivity \\[4pt]
    \textbf{Persistent (\$9/mo)} & 2 vCPU & 16\,GB & Never sleeps \\[4pt]
    GPU A10G (\$1.05/hr) & GPU & 24\,GB & On-demand; only for training \\

\end{longtable}

The \textbf{Persistent CPU tier at \$9/month} is the correct choice.
Training is performed separately and models are loaded from HF Hub
at startup; GPU is not required for inference at this scale.

\section{CDN, SSL, and DNS: Cloudflare (Free Tier)}

Cloudflare's free tier provides everything required:

\begin{itemize}[leftmargin=1.5em]
    \item Automatic HTTPS with SSL certificate management
    \item Global CDN for Vite-compiled static assets
          (content-hashed filenames allow aggressive caching)
    \item Web Application Firewall with basic OWASP rule sets
    \item DDoS protection
    \item DNS management with \textless 1-minute propagation
    \item No cost for the coverage needed at thesis scale
\end{itemize}

\section{Object Storage: Cloudflare R2}

Cloudflare R2 replaces the local disk for file storage:

\begin{itemize}[leftmargin=1.5em]
    \item S3-compatible API: works with Laravel's existing
          \texttt{league/flysystem-aws-s3-v3} driver with no code changes
    \item Free tier: 10\,GB storage, 10 million read operations/month
    \item Zero egress fees (unlike AWS S3)
    \item Served through Cloudflare's global network
\end{itemize}

\section{Cache and Sessions: Database Driver}

At thesis demo scale (low concurrent users), the \texttt{database}
driver for both sessions and cache avoids introducing Redis as a
dependency:

\begin{lstlisting}[caption={Production .env session/cache settings}]
SESSION_DRIVER=database
CACHE_STORE=database
\end{lstlisting}

This requires two additional Laravel migrations
(\texttt{php artisan session:table} and
\texttt{php artisan cache:table}).

If Redis is desired in future, \textbf{Upstash Redis} is the correct
choice: serverless Redis with 10,000 free requests/day, pay-per-use
above that.

\section{Queue Driver: Database}

Since no \texttt{app/Jobs/} directory exists and all current work is
synchronous:

\begin{lstlisting}[caption={Production .env queue settings}]
QUEUE_CONNECTION=database
\end{lstlisting}

Jobs are stored in the PostgreSQL \texttt{jobs} table. A Railway
worker process runs:

\begin{lstlisting}
php artisan queue:work --sleep=3 --tries=3 --timeout=60
\end{lstlisting}

This is sufficient for thesis demo scale without adding Redis
infrastructure.

\section{Summary Technology Decisions}

\begin{longtable}{L{3cm} L{3.5cm} L{5.5cm}}
    \toprule
    \textbf{Decision} & \textbf{Choice} & \textbf{Rationale} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Decision} & \textbf{Choice} & \textbf{Rationale} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Laravel host & Railway & Multi-process, Docker-native, affordable \\[4pt]
    AI service & HF Spaces (persistent) & Already deployed, HF Hub sync \\[4pt]
    Database & Supabase & Already configured, SSL, pooling \\[4pt]
    Object storage & Cloudflare R2 & Free egress, S3-compatible \\[4pt]
    CDN + SSL & Cloudflare & Free, global, WAF included \\[4pt]
    CI/CD & GitHub Actions & Free tier sufficient \\[4pt]
    Queue & Database driver & No Redis overhead at this scale \\[4pt]
    Session/Cache & Database driver & Same reasoning \\[4pt]
    Error tracking & Sentry & Best free tier, Laravel + FastAPI plugins \\[4pt]
    Monitoring & UptimeRobot & Free, sufficient for thesis alerting \\[4pt]
    Container registry & GHCR & Free, integrated with Actions \\[4pt]
    Reverse proxy & Nginx (in container) & Standard PHP-FPM companion \\

\end{longtable}

% ════════════════════════════════════════════════════════════
\chapter{Security Architecture Recommendations}
\label{ch:security}
% ════════════════════════════════════════════════════════════

\section{Authentication Security}

The system implements dual-guard authentication:
\textbf{JWT} (\texttt{tymon/jwt-auth}) for API consumers and
\textbf{session-based} authentication for the Inertia.js web
interface. The \texttt{RefreshTokenFamily} model implements token
rotation and replay attack protection. RBAC is handled by
\texttt{spatie/laravel-permission}.

\begin{itemize}[leftmargin=1.5em]
    \item \texttt{JWT\_SECRET} must be a 256-bit random secret injected
          via Railway environment secrets, never in any committed file.
    \item \texttt{APP\_KEY} must be injected as a secret.
    \item JWT \texttt{TTL=60} minutes is reasonable; ensure refresh
          token expiry is configured appropriately.
    \item \texttt{SANCTUM\_STATEFUL\_DOMAINS} must be set to the
          production domain for session-based auth to function correctly.
\end{itemize}

\section{Network Security}

\subsection{Transport Layer}

All traffic must be HTTPS. Cloudflare handles SSL termination at the
edge. The origin server (Railway) communicates internally via HTTP
since Cloudflare encrypts the public edge.

For stricter security, enable Cloudflare's \textbf{Full (strict)} SSL
mode, which requires a valid origin certificate. Cloudflare provides
free Origin CA certificates for this purpose.

\subsection{Inter-Service Authentication}

The AI service validates every request using an \texttt{X-Internal-Key}
header (implemented in \texttt{app/main.py} middleware). This shared
secret must:

\begin{itemize}[leftmargin=1.5em]
    \item Be a non-guessable string of at least 32 random characters.
    \item Be stored as a secret in both Railway (\texttt{AI\_SERVICE\_TOKEN})
          and HuggingFace Spaces (\texttt{INTERNAL\_API\_KEY}).
    \item Match exactly between both services.
    \item Be rotated and never committed to any file.
\end{itemize}

\subsection{Database Security}

\begin{itemize}[leftmargin=1.5em]
    \item Supabase enforces \texttt{DB\_SSLMODE=require} — already correct.
    \item The database port must not be exposed publicly.
    \item Use Supabase's Row Level Security for multi-tenant isolation
          where applicable.
\end{itemize}

\section{Application Security}

\subsection{HTTP Security Headers}

The Nginx configuration must include:

\begin{lstlisting}[caption={Required Nginx security headers}]
add_header X-Content-Type-Options "nosniff";
add_header X-Frame-Options "SAMEORIGIN";
add_header X-XSS-Protection "1; mode=block";
add_header Referrer-Policy "strict-origin-when-cross-origin";
add_header Permissions-Policy "camera=(), microphone=(), geolocation=()";
\end{lstlisting}

A \texttt{Content-Security-Policy} header requires careful configuration
with Vite's content-hashed asset filenames.

\subsection{CSRF Protection}

Laravel's CSRF protection is already built in for web routes.
The session domain must be correctly configured in production to match
the public domain.

\subsection{Rate Limiting}

Laravel's \texttt{throttle} middleware must be applied to API route
groups in \texttt{routes/api.php} before launch:

\begin{lstlisting}[caption={API rate limiting (routes/api.php)}]
Route::middleware(['auth:api', 'throttle:60,1'])->group(function () {
    // ... all API routes
});
\end{lstlisting}

\subsection{Debug Mode}

\texttt{APP\_DEBUG} \textbf{must be \texttt{false}} in production.
Debug mode exposes stack traces, environment variables, and database
queries directly to the browser.

\section{Compliance and GDPR}

The system already implements GDPR infrastructure:
\texttt{Consent}, \texttt{DataSubjectRequest}, and
\texttt{AdverseActionNotice} models are present. For production:

\begin{itemize}[leftmargin=1.5em]
    \item Erasure requests (\texttt{POST /v1/me/privacy/erasure-requests})
          must actually delete or anonymise data, not merely soft-delete.
    \item The audit log (\texttt{owen-it/laravel-auditing}) must be
          append-only and tamper-evident.
    \item Consent must be captured before any valuation data is
          processed.
    \item The Python service's \texttt{structlog} output must not log
          PII or query parameters.
\end{itemize}

% ════════════════════════════════════════════════════════════
\chapter{Deployment Strategy}
\label{ch:deployment}
% ════════════════════════════════════════════════════════════

\section{Environment Promotion Path}

For a thesis project, the minimum viable promotion path is:

\begin{center}
\texttt{local (development)} \(\longrightarrow\) \texttt{production (Railway)}
\end{center}

With a separate \texttt{.env} for each environment. No code changes
are required between environments — only environment variable
differences.

\section{Deployment Topology}

All three Laravel processes run from the same Docker image with
different start commands within one Railway project:

\begin{lstlisting}[caption={Railway project service layout}]
Railway Project: ethio-sme
+-- Service: web
|   Start command: nginx (via PHP-FPM + Nginx entrypoint)
|   Port: 80
|   Health check: GET /up
|
+-- Service: worker
|   Start command: php artisan queue:work --sleep=3 --tries=3
|   No port exposed
|
+-- Service: scheduler
    Start command: php artisan schedule:run
    Or: Railway native cron (every minute)
\end{lstlisting}

All three services share environment variables and connect to the
same Supabase database.

\section{Production Dockerfile Requirements}

The current \texttt{Dockerfile} (based on Laravel Sail) must be
replaced. The new production image requires:

\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{Base image:} \texttt{php:8.3-fpm-alpine}
    \item \textbf{Web server:} Nginx installed in same container
    \item \textbf{PHP extensions:} \texttt{pgsql}, \texttt{pdo\_pgsql},
          \texttt{redis}, \texttt{gd}, \texttt{zip}, \texttt{bcmath},
          \texttt{intl}
    \item \textbf{Composer dependencies:} \texttt{--no-dev --optimize-autoloader}
    \item \textbf{Frontend build:} \texttt{npm run build} executed during image build
    \item \textbf{Laravel caches:} \texttt{config:cache}, \texttt{route:cache},
          \texttt{view:cache} applied in entrypoint
    \item \textbf{Non-root user:} Container runs as unprivileged user
    \item \textbf{Excluded:} Xdebug, Vite dev server, Sail tooling
\end{enumerate}

\section{Container Entrypoint (Web)}

On container start for the web service:

\begin{lstlisting}[caption={Production entrypoint sequence}]
1. php artisan migrate --force
2. php artisan config:cache
3. php artisan route:cache
4. php artisan view:cache
5. Start PHP-FPM (background)
6. Start Nginx (foreground)
\end{lstlisting}

\section{Deployment Flow}

\begin{lstlisting}[caption={End-to-end deployment flow}]
git push origin main
     |
     v
GitHub Actions CI triggered
     |
     +-- Run php artisan test (SQLite in-memory)
     +-- Run vendor/bin/pint --test
     +-- Run npm run build (verify frontend compiles)
     |
     v
CI passes
     |
     v
Build Docker image
     |
     v
Push to GitHub Container Registry (GHCR)
     |
     v
Railway pulls new image, rolling deploy
     |
     v
Post-deploy hook: php artisan migrate --force
     |
     v
Smoke test: curl /up --> assert 200
\end{lstlisting}

% ════════════════════════════════════════════════════════════
\chapter{CI/CD Strategy}
\label{ch:cicd}
% ════════════════════════════════════════════════════════════

The project currently has no \texttt{.github/workflows/} directory.
Two GitHub Actions workflows are required.

\section{Workflow 1: Continuous Integration}

\begin{lstlisting}[caption={CI workflow (every push and pull request to main)}]
Trigger: push (any branch), pull_request (target: main)

Steps:
  1.  actions/checkout@v4
  2.  Setup PHP 8.3 (shivammathur/setup-php)
  3.  Cache Composer dependencies (actions/cache)
  4.  composer install --no-interaction --prefer-dist
  5.  Setup Node.js 20 (actions/setup-node)
  6.  Cache npm dependencies (actions/cache)
  7.  npm ci
  8.  npm run build            (verify TypeScript compiles)
  9.  cp .env.testing .env
  10. php artisan config:clear
  11. php artisan test          (SQLite in-memory per phpunit.xml)
  12. vendor/bin/pint --test    (lint check, no changes)
\end{lstlisting}

\section{Workflow 2: Continuous Deployment}

\begin{lstlisting}[caption={Deploy workflow (push to main only, after CI passes)}]
Trigger: workflow_run (CI workflow: completed + success)
         (or: push to main with CI as required status check)

Steps:
  1. actions/checkout@v4
  2. Log in to GHCR (docker/login-action)
  3. Build Docker image (docker/build-push-action)
  4. Push to ghcr.io/<owner>/ethio-sme-backend:latest
  5. Trigger Railway deploy via Railway CLI
     (railway up --service=web --detach)
  6. Run post-deploy migration via Railway CLI
     (railway run --service=web php artisan migrate --force)
  7. Smoke test: curl https://<domain>/up
     Assert HTTP 200 response
\end{lstlisting}

\section{AI Service CI/CD}

The Python FastAPI service (\texttt{Ethiopian-SME-AI-Service} repo)
requires a separate, simpler workflow:

\begin{lstlisting}[caption={AI service deploy workflow}]
Trigger: push to main

Steps:
  1. actions/checkout@v4
  2. Setup Python 3.11
  3. pip install -r requirements.txt
  4. pytest tests/          (run all tests)
  5. Push to HuggingFace Spaces via git:
     git push https://huggingface.co/spaces/<org>/... main
     (authenticated via HF_TOKEN secret)
\end{lstlisting}

\section{Branch Strategy}

\begin{lstlisting}
main          --> auto-deploys to production
feature/*     --> CI runs, no deploy
\end{lstlisting}

For a thesis project, trunk-based development is appropriate. No
need for \texttt{develop}, \texttt{staging}, or \texttt{release}
branches.

% ════════════════════════════════════════════════════════════
\chapter{Monitoring and Logging Strategy}
\label{ch:monitoring}
% ════════════════════════════════════════════════════════════

\section{Error Tracking: Sentry}

\textbf{Sentry} (free tier: 5,000 errors/month) must be integrated
in both services.

\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Laravel:} Install \texttt{sentry/sentry-laravel}.
          Configure \texttt{SENTRY\_DSN} as environment variable.
          Captures unhandled exceptions, slow queries, and deployments.
    \item \textbf{Python:} Install \texttt{sentry-sdk[fastapi]}.
          Configure \texttt{SENTRY\_DSN}. Captures unhandled exceptions
          and transaction traces.
\end{itemize}

Sentry is the highest-leverage single observability addition for a
thesis project. It means failures are known before the supervisor
sees them.

\section{Application Performance Monitoring}

Sentry's built-in performance monitoring (5,000 transactions/month
on the free tier) captures:

\begin{itemize}[leftmargin=1.5em]
    \item Slow database queries (with query plans)
    \item Slow external HTTP calls (AI service latency distribution)
    \item Request duration histograms by endpoint
    \item Memory usage spikes
\end{itemize}

\section{Logging Configuration}

\subsection{Laravel}

\begin{lstlisting}[caption={Production logging .env}]
LOG_CHANNEL=stderr
LOG_LEVEL=warning
\end{lstlisting}

Railway captures all \texttt{stderr}/\texttt{stdout} and provides a
log viewer in the dashboard with 7-day retention. Warning level
suppresses debug and info noise while preserving actionable messages.

\subsection{Python AI Service}

The AI service already uses \texttt{structlog} for structured JSON
logging to \texttt{stdout}. HuggingFace Spaces captures this in the
Space logs panel. Log level should remain at \texttt{info} — the
structured format is parseable and each request is already logged
with \texttt{request\_id}, path, status, and latency.

\section{Uptime Monitoring}

\textbf{UptimeRobot} (free tier: 50 monitors, 5-minute check interval)
must monitor:

\begin{longtable}{L{5cm} L{4cm} L{4cm}}
    \toprule
    \textbf{Endpoint} & \textbf{Service} & \textbf{Alert on} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Endpoint} & \textbf{Service} & \textbf{Alert on} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    \texttt{https://domain.com/up} & Laravel & Non-200 response \\[4pt]
    \texttt{https://...hf.space/health} & FastAPI AI service &
        Non-200 or model \texttt{loaded: false} \\

\end{longtable}

Alert delivery: email and/or Telegram. Uptime monitoring catches a
sleeping HuggingFace Space before the demo does.

\section{Business Metrics}

No custom metrics dashboard is required for a thesis project. The
existing \texttt{FairnessAudit}, \texttt{DriftMetric}, and
\texttt{AiEvaluationLog} database tables serve as the application's
own telemetry. The Super Admin dashboard already surfaces these.

% ════════════════════════════════════════════════════════════
\chapter{Backup and Disaster Recovery}
\label{ch:backup}
% ════════════════════════════════════════════════════════════

\section{Database Backups}

\subsection{Supabase Automated Backups}

\begin{longtable}{L{3.5cm} L{3.5cm} L{5cm}}
    \toprule
    \textbf{Supabase Tier} & \textbf{Backup Retention} & \textbf{Features} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Supabase Tier} & \textbf{Backup Retention} & \textbf{Features} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Free & 7 days & Daily snapshots only \\[4pt]
    Pro (\$25/mo) & 30 days & Daily snapshots + PITR (1-day window) \\

\end{longtable}

\subsection{Manual Pre-Demo Backup}

Before major demonstrations, a manual \texttt{pg\_dump} backup must
be taken and stored offline:

\begin{lstlisting}[caption={Manual database backup command}]
pg_dump "postgresql://postgres.<ref>:<password>@aws-1-eu-central-1.pooler.supabase.com:5432/postgres" \
    --format=custom \
    --file="backup-$(date +%Y%m%d-%H%M).dump"
\end{lstlisting}

Store in a private GitHub repository or Cloudflare R2 bucket.

\section{Model Artifact Backups}

The \texttt{storage\_service.py} already syncs trained models to
HuggingFace Hub. This provides:

\begin{itemize}[leftmargin=1.5em]
    \item Versioned artifact store with git-based history
    \item \texttt{versions.json} tracks current model versions and
          training dates
    \item Rollback to any previous version by updating \texttt{versions.json}
\end{itemize}

No additional model backup infrastructure is required.

\section{Application Code}

Git history on GitHub is the primary code backup. No additional code
backup is necessary.

\section{Disaster Recovery Scenarios}

\begin{longtable}{L{3.5cm} C{1.8cm} C{1.8cm} L{4cm}}
    \toprule
    \textbf{Scenario} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Recovery Action} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Scenario} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Recovery Action} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Railway deploy fails &
    \textcolor{warningorange}{Low} &
    \textcolor{dangerred}{High} &
    Railway dashboard: 1-click rollback to previous deployment \\[6pt]

    Database corruption &
    \textcolor{successgreen}{Very Low} &
    \textcolor{dangerred}{High} &
    Restore from Supabase daily backup \\[6pt]

    HF Space sleeping during demo &
    \textcolor{dangerred}{High (free tier)} &
    \textcolor{dangerred}{High} &
    Upgrade to persistent; keep-alive pings every 30 min \\[6pt]

    ML models not loaded &
    \textcolor{warningorange}{Medium} &
    \textcolor{dangerred}{High} &
    \texttt{GET /health} check in CI; retry logic in \texttt{ValuationService} \\[6pt]

    Valuation timeout &
    \textcolor{warningorange}{Medium} &
    \textcolor{warningorange}{Medium} &
    Increase \texttt{AI\_SERVICE\_TIMEOUT} to 120 seconds \\[6pt]

    Environment variable missing &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    Validate all required vars in entrypoint script \\[6pt]

    Supabase paused (inactivity) &
    \textcolor{warningorange}{Medium} &
    \textcolor{dangerred}{High} &
    Hourly DB ping via cron; or upgrade to Pro \\[6pt]

    Model artifacts corrupted &
    \textcolor{successgreen}{Low} &
    \textcolor{warningorange}{Medium} &
    Reload previous version from HF Hub via \texttt{versions.json} \\[6pt]

    Contract version mismatch &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    Pin \texttt{AI\_SERVICE\_CONTRACT\_VERSION=v1} in both services \\

\end{longtable}

\textbf{RTO (Recovery Time Objective) for thesis:} 15 minutes for all
scenarios except database restore (up to 24 hours for full restore;
partial restore faster).

% ════════════════════════════════════════════════════════════
\chapter{Estimated Hosting Cost}
\label{ch:cost}
% ════════════════════════════════════════════════════════════

\section{Baseline (Thesis Demo) Cost}

\begin{longtable}{L{5cm} L{3.5cm} R{3cm}}
    \toprule
    \textbf{Service} & \textbf{Tier} & \textbf{Monthly Cost} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Service} & \textbf{Tier} & \textbf{Monthly Cost} \\
    \midrule
    \endhead
    \midrule
    \textbf{Total} & & \textbf{\$9 -- \$14 / month} \\
    \bottomrule
    \endlastfoot

    Railway (web + worker + cron) & Starter (\$5 credit/mo free) & \$0 -- 5 \\[4pt]
    Supabase PostgreSQL & Free tier & \$0 \\[4pt]
    HuggingFace Spaces (AI) & Persistent CPU & \$9 \\[4pt]
    Cloudflare (CDN + SSL + DNS) & Free & \$0 \\[4pt]
    Cloudflare R2 (object storage) & Free (10\,GB) & \$0 \\[4pt]
    GitHub Actions (CI/CD) & Free (2,000 min/mo) & \$0 \\[4pt]
    Sentry (error tracking) & Free (5K events) & \$0 \\[4pt]
    UptimeRobot (monitoring) & Free & \$0 \\[4pt]
    Domain name (optional) & Namecheap annual & \$1 \\

\end{longtable}

\begin{infobox}[Cost Summary]
    The full production stack — with zero infrastructure compromises
    for a thesis demonstration — costs \textbf{\$9 to \$14 per month}.
    If the HuggingFace free tier is kept warm via pings and monitored
    carefully, the cost can be reduced to \textbf{\$0 to \$5 per month}.
\end{infobox}

\section{Upgrade Triggers and Costs}

\begin{longtable}{L{4.5cm} L{4cm} R{3.5cm}}
    \toprule
    \textbf{Trigger} & \textbf{Upgrade} & \textbf{Additional Cost} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Trigger} & \textbf{Upgrade} & \textbf{Additional Cost} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    Supabase pausing repeatedly & Supabase Pro & +\$25/mo \\[4pt]
    Railway CPU exhaustion & Railway Pro & +\$5 -- 20/mo \\[4pt]
    AI inference too slow & HF Spaces GPU A10G & +\$1.05/hr (on-demand) \\[4pt]
    File storage \textgreater\ 10\,GB & R2 paid tier & +\$0.015/GB \\[4pt]
    High concurrent users & Redis (Upstash) & +\$0 -- 10/mo \\

\end{longtable}

% ════════════════════════════════════════════════════════════
\chapter{Step-by-Step Deployment Roadmap}
\label{ch:roadmap}
% ════════════════════════════════════════════════════════════

\section{Phase 1: Production-Ready Application (Week 1)}

\textbf{Goal:} Make the Laravel codebase deployable to any Docker-based
platform without localhost assumptions.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}]

    \item \textbf{Replace Dockerfile} \\
    Write a production Dockerfile using \texttt{php:8.3-fpm-alpine} +
    Nginx. Remove the Sail base image. Ensure \texttt{npm run build}
    runs during the Docker build stage. Multi-stage build recommended
    (Node.js stage for assets, PHP stage for application).

    \item \textbf{Replace docker-entrypoint.sh} \\
    New entrypoint for the web service: run migrations, cache config,
    routes, and views, then start PHP-FPM and Nginx. No Vite dev server.

    \item \textbf{Update .env for production} \\
    Set: \texttt{APP\_ENV=production}, \texttt{APP\_DEBUG=false},
    \texttt{LOG\_CHANNEL=stderr}, \texttt{SESSION\_DRIVER=database},
    \texttt{CACHE\_STORE=database}, \texttt{QUEUE\_CONNECTION=database},
    \texttt{FILESYSTEM\_DISK=s3}.

    \item \textbf{Add database migrations} \\
    Run \texttt{php artisan session:table}, \texttt{cache:table},
    \texttt{queue:table} to generate the required migrations.
    Commit the generated migration files.

    \item \textbf{Verify health endpoint} \\
    Confirm \texttt{GET /up} returns 200 (Laravel 11+ default route).
    If not present, add it to \texttt{routes/web.php}.

    \item \textbf{Switch logging to stderr} \\
    Update \texttt{config/logging.php} to add a \texttt{stderr}
    channel and set it as the default in the production environment.

\end{enumerate}

\section{Phase 2: Infrastructure Setup (Week 1--2)}

\textbf{Goal:} Provision all cloud services and connect them.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=7]

    \item \textbf{Domain registration (optional)} \\
    Register a domain via Namecheap (\textasciitilde\$10/year) or use
    the free Railway subdomain (e.g., \texttt{ethio-sme.up.railway.app})
    for the thesis.

    \item \textbf{Cloudflare setup} \\
    Add domain to Cloudflare. Update nameservers at the registrar.
    Enable proxy (orange cloud) for all A/CNAME records. This activates
    SSL, CDN, and WAF automatically.

    \item \textbf{Railway project creation} \\
    Create a Railway project. Connect the GitHub repository. Add all
    environment variables (all values from the production \texttt{.env}).
    Never commit secrets.

    \item \textbf{Railway services} \\
    Create three services from the same Docker image:
    \texttt{web} (Nginx + PHP-FPM), \texttt{worker}
    (\texttt{queue:work}), \texttt{scheduler}
    (\texttt{schedule:run} cron).

    \item \textbf{Cloudflare R2 bucket} \\
    Create an R2 bucket. Generate API access key. Add
    \texttt{AWS\_ACCESS\_KEY\_ID}, \texttt{AWS\_SECRET\_ACCESS\_KEY},
    \texttt{AWS\_BUCKET}, \texttt{AWS\_ENDPOINT} (R2-specific endpoint)
    to Railway environment secrets.

    \item \textbf{HuggingFace Space upgrade} \\
    Upgrade the existing Space to the persistent CPU tier (\$9/mo).
    Verify the Space URL is accessible and does not suspend.

\end{enumerate}

\section{Phase 3: CI/CD Pipeline (Week 2)}

\textbf{Goal:} Every push to \texttt{main} tests and deploys automatically.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=13]

    \item \textbf{GitHub Actions CI workflow} \\
    Create \texttt{.github/workflows/ci.yml} that runs tests and
    linting on every push to any branch.

    \item \textbf{GitHub Actions Deploy workflow} \\
    Create \texttt{.github/workflows/deploy.yml} that builds and
    pushes the Docker image to GHCR, then triggers Railway deploy.

    \item \textbf{Test the full pipeline} \\
    Push a trivial change to \texttt{main}. Verify CI passes, deploy
    triggers, Railway pulls the new image, and the health check passes.

\end{enumerate}

\section{Phase 4: Security Hardening (Week 2)}

\textbf{Goal:} Apply all production security controls.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=16]

    \item \textbf{Rotate all secrets} \\
    Generate new \texttt{APP\_KEY} (\texttt{php artisan key:generate
    --show}), \texttt{JWT\_SECRET}, \texttt{AI\_SERVICE\_TOKEN},
    \texttt{INTERNAL\_API\_KEY}. Store in Railway secrets. Keep an
    offline copy in a password manager.

    \item \textbf{Cloudflare WAF rules} \\
    Enable managed OWASP Core Rule Set on Cloudflare's free WAF.
    Block known bad user agents.

    \item \textbf{Rate limiting} \\
    Add \texttt{throttle:60,1} middleware to API route groups in
    \texttt{routes/api.php}.

    \item \textbf{Security headers in Nginx} \\
    Add all required HTTP security headers to the production Nginx
    configuration file.

\end{enumerate}

\section{Phase 5: Monitoring and Observability (Week 2--3)}

\textbf{Goal:} Know about failures before users do.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=20]

    \item \textbf{Sentry integration} \\
    Add \texttt{sentry/sentry-laravel} to \texttt{composer.json}.
    Configure \texttt{SENTRY\_DSN} in Railway secrets.
    Add \texttt{sentry-sdk[fastapi]} to Python \texttt{requirements.txt}.
    Configure \texttt{SENTRY\_DSN} in HuggingFace Spaces secrets.

    \item \textbf{UptimeRobot monitors} \\
    Add monitors for \texttt{/up} (Laravel) and \texttt{/health}
    (AI service). Configure email or Telegram alerts.

    \item \textbf{HuggingFace Space keep-alive} \\
    Add a Railway cron job (scheduled service) that sends a
    \texttt{GET} request to
    \texttt{https://\ldots{}hf.space/health} every 30 minutes.
    This prevents the Space from sleeping between thesis sessions
    even on the free tier.

\end{enumerate}

\section{Phase 6: Demo Data and Smoke Testing (Week 3)}

\textbf{Goal:} Confirm all features work end-to-end on the production URL.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=23]

    \item \textbf{Run migrations on production} \\
    \texttt{php artisan migrate --force} via Railway CLI or
    post-deploy hook. Confirm all tables exist.

    \item \textbf{Seed demo data} \\
    \texttt{php artisan db:seed --class=DevDemoSeeder} as documented
    in \texttt{CLAUDE.md}. Verify realistic borrower, application,
    and valuation records are created.

    \item \textbf{End-to-end smoke test} \\
    Manually walk through every major user flow:
    \begin{itemize}[leftmargin=1em]
        \item Register a borrower (SME owner)
        \item Submit a loan application
        \item Complete psychometric assessment
        \item Trigger valuation (AI pipeline)
        \item Evaluate as a lender
        \item Review audit logs as Super Admin
    \end{itemize}

    \item \textbf{AI service warm-up} \\
    Trigger a training job (\texttt{POST /train}) after a fresh
    deploy. Confirm \texttt{GET /health} returns all models as
    \texttt{loaded: true} before declaring the service ready.

\end{enumerate}

\section{Phase 7: Pre-Defense Hardening (Week Before Defense)}

\textbf{Goal:} Eliminate all variables that could cause a demo failure.

\begin{enumerate}[leftmargin=1.5em, label=\textbf{\arabic*.}, start=27]

    \item \textbf{Database backup} \\
    Run manual \texttt{pg\_dump} backup. Store offline.

    \item \textbf{Feature freeze} \\
    No new features or refactors during the week before the defense.
    Only critical bug fixes allowed.

    \item \textbf{Demo rehearsal} \\
    Run the complete thesis demo scenario at least twice on the
    production URL, including model training and valuation.

    \item \textbf{Document failover plan} \\
    Write and keep accessible: exact steps for the three most likely
    failures:
    \begin{itemize}[leftmargin=1em]
        \item \textbf{App is down:} Railway dashboard → Deployments → Rollback
        \item \textbf{AI service sleeping:} HuggingFace Spaces dashboard → Restart
        \item \textbf{Database paused:} Supabase dashboard → Project → Resume
    \end{itemize}

\end{enumerate}

% ════════════════════════════════════════════════════════════
\chapter{Risk Register}
\label{ch:risks}
% ════════════════════════════════════════════════════════════

\begin{longtable}{L{3.5cm} C{2cm} C{1.8cm} L{5cm}}
    \toprule
    \textbf{Risk} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Mitigation} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Risk} & \textbf{Likelihood} & \textbf{Impact} & \textbf{Mitigation} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    HuggingFace Space sleeps during demo &
    \textcolor{dangerred}{High} &
    \textcolor{dangerred}{High} &
    Upgrade to persistent tier; 30-min keep-alive ping \\[6pt]

    Supabase pauses for inactivity &
    \textcolor{warningorange}{Medium} &
    \textcolor{dangerred}{High} &
    Hourly DB ping cron; or Supabase Pro \\[6pt]

    ML models not loaded at startup &
    \textcolor{warningorange}{Medium} &
    \textcolor{dangerred}{High} &
    \texttt{GET /health} check in CI smoke test; retry logic \\[6pt]

    Valuation timeout (DeepAR slow) &
    \textcolor{warningorange}{Medium} &
    \textcolor{warningorange}{Medium} &
    Increase \texttt{AI\_SERVICE\_TIMEOUT} to 120 seconds \\[6pt]

    Railway cold start &
    \textcolor{successgreen}{Low} &
    \textcolor{warningorange}{Medium} &
    UptimeRobot pings keep container warm \\[6pt]

    Env variable missing on deploy &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    Entrypoint validation; offline backup in password manager \\[6pt]

    Schema drift between services &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    Laravel migrations are source of truth; AI uses raw SQL \\[6pt]

    SHAP memory spike &
    \textcolor{successgreen}{Low} &
    \textcolor{warningorange}{Medium} &
    HF persistent CPU has 16\,GB RAM; monitor during load \\[6pt]

    Contract version mismatch (v1) &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    Pin contract version in both services; version-check on \texttt{/health} \\[6pt]

    Secret leaked in logs &
    \textcolor{successgreen}{Low} &
    \textcolor{dangerred}{High} &
    \texttt{LOG\_LEVEL=warning}; structlog no-PII policy; Sentry scrubbing \\

\end{longtable}

% ════════════════════════════════════════════════════════════
\chapter{What Should Remain Unchanged}
\label{ch:unchanged}
% ════════════════════════════════════════════════════════════

The following are well-designed elements of the codebase that
\textbf{should not be refactored before deployment}. Deployment does
not change business logic or internal architecture.

\begin{itemize}[leftmargin=1.5em]

    \item \textbf{Domain-Driven Design structure} --- The 10-domain
    layout in \texttt{app/Domain/} is clean, well-organised, and
    deployment is orthogonal to it.

    \item \textbf{Supabase PostgreSQL schema} --- Already configured
    and working. Migrations are the source of truth.

    \item \textbf{HuggingFace Hub model storage} --- The
    \texttt{storage\_service.py} versioning system is well-designed.

    \item \textbf{Spatie DTOs} --- Type safety via
    \texttt{spatie/laravel-data} adds no deployment overhead.

    \item \textbf{RBAC via spatie/laravel-permission} --- Role
    middleware aliases are already correctly configured for both guards.

    \item \textbf{Audit trail via owen-it/laravel-auditing} --- Passive
    observer; no changes needed.

    \item \textbf{JWT dual-guard authentication} --- The dual-guard
    pattern is correct for Inertia + REST API.

    \item \textbf{Python service Pydantic schemas} --- The v1 contract
    is stable.

    \item \textbf{RefreshTokenFamily replay protection} --- Security
    feature; must not be removed or simplified.

    \item \textbf{AI service middleware chain} --- Key validation,
    request logging, and exception handling in \texttt{app/main.py}
    are correct and must not be modified.

\end{itemize}

% ════════════════════════════════════════════════════════════
\appendix
% ════════════════════════════════════════════════════════════

\chapter{Environment Variable Reference}
\label{app:envvars}

The following table lists all environment variables required for the
production Laravel deployment, grouped by concern.

\begin{longtable}{L{5.5cm} L{6.5cm}}
    \toprule
    \textbf{Variable} & \textbf{Production Value} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Variable} & \textbf{Production Value} \\
    \midrule
    \endhead
    \bottomrule
    \endlastfoot

    \multicolumn{2}{l}{\textbf{Application}} \\[4pt]
    \texttt{APP\_NAME} & EthioSME Valuation \\
    \texttt{APP\_ENV} & production \\
    \texttt{APP\_DEBUG} & false \\
    \texttt{APP\_KEY} & (generate: \texttt{php artisan key:generate --show}) \\
    \texttt{APP\_URL} & \texttt{https://your-domain.com} \\[8pt]

    \multicolumn{2}{l}{\textbf{Database (Supabase)}} \\[4pt]
    \texttt{DB\_CONNECTION} & pgsql \\
    \texttt{DB\_HOST} & (Supabase pooler host) \\
    \texttt{DB\_PORT} & 5432 \\
    \texttt{DB\_DATABASE} & postgres \\
    \texttt{DB\_USERNAME} & postgres.\textit{ref} \\
    \texttt{DB\_PASSWORD} & (Supabase DB password) \\
    \texttt{DB\_SSLMODE} & require \\[8pt]

    \multicolumn{2}{l}{\textbf{Session, Cache, Queue}} \\[4pt]
    \texttt{SESSION\_DRIVER} & database \\
    \texttt{CACHE\_STORE} & database \\
    \texttt{QUEUE\_CONNECTION} & database \\[8pt]

    \multicolumn{2}{l}{\textbf{File Storage (Cloudflare R2)}} \\[4pt]
    \texttt{FILESYSTEM\_DISK} & s3 \\
    \texttt{AWS\_ACCESS\_KEY\_ID} & (R2 access key) \\
    \texttt{AWS\_SECRET\_ACCESS\_KEY} & (R2 secret key) \\
    \texttt{AWS\_BUCKET} & (R2 bucket name) \\
    \texttt{AWS\_ENDPOINT} & (R2 S3-compatible endpoint URL) \\
    \texttt{AWS\_DEFAULT\_REGION} & auto \\[8pt]

    \multicolumn{2}{l}{\textbf{Authentication}} \\[4pt]
    \texttt{JWT\_SECRET} & (256-bit random secret) \\
    \texttt{JWT\_TTL} & 60 \\[8pt]

    \multicolumn{2}{l}{\textbf{AI Service}} \\[4pt]
    \texttt{AI\_SERVICE\_URL} & (HuggingFace Spaces URL) \\
    \texttt{AI\_SERVICE\_TOKEN} & (32+ char shared secret) \\
    \texttt{AI\_SERVICE\_TIMEOUT} & 120 \\
    \texttt{AI\_SERVICE\_CONTRACT\_VERSION} & v1 \\
    \texttt{AI\_SERVICE\_FALLBACK\_ENABLED} & true \\[8pt]

    \multicolumn{2}{l}{\textbf{Logging and Monitoring}} \\[4pt]
    \texttt{LOG\_CHANNEL} & stderr \\
    \texttt{LOG\_LEVEL} & warning \\
    \texttt{SENTRY\_DSN} & (Sentry project DSN) \\[8pt]

    \multicolumn{2}{l}{\textbf{Mail (Resend)}} \\[4pt]
    \texttt{MAIL\_MAILER} & resend \\
    \texttt{RESEND\_API\_KEY} & (Resend API key) \\
    \texttt{MAIL\_FROM\_ADDRESS} & noreply@your-domain.com \\

\end{longtable}

\chapter{Deployment Checklist}
\label{app:checklist}

Use this checklist immediately before and after each production
deployment.

\subsection*{Pre-Deployment}

\begin{itemize}[leftmargin=1.5em, label=$\square$]
    \item All environment variables set in Railway secrets
    \item \texttt{APP\_DEBUG=false} confirmed
    \item \texttt{APP\_ENV=production} confirmed
    \item CI pipeline passes (tests + lint)
    \item Docker image builds successfully locally
    \item HuggingFace Space is awake (\texttt{GET /health} returns 200)
    \item Supabase project is active (not paused)
\end{itemize}

\subsection*{Post-Deployment}

\begin{itemize}[leftmargin=1.5em, label=$\square$]
    \item \texttt{GET /up} returns 200
    \item \texttt{php artisan migrate --force} completed successfully
    \item No errors in Railway deployment logs
    \item \texttt{GET https://...hf.space/health} returns all models
          as \texttt{loaded: true}
    \item Sentry receives a test exception (or no new alerts)
    \item UptimeRobot shows all monitors green
    \item Login flow works end-to-end
    \item Valuation pipeline completes without timeout
\end{itemize}

\end{document}
